---
title: "Manipulating 2018 Cal Fire damage data"
author: "Hanna Buechi"
date: "11/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the data that Steve XXX at Cal Fire sent to me and Dick. It includes information about *each house* (!!!), down to the street number, that was damaged or destroyed in wildfires between 2013 and 2018. I just need the 2018 stuff.

```{r data_packages}

library(tidyverse)
library(lubridate)

recent_damage <- read_csv(
  "DINS2013_2018_attributes.csv",
  col_types = cols(
    REPORTED = col_character(),
    BATTALION = col_character(),
    FIRENAME = col_character()
  ))

recent_damage$INCIDENTST <- as.Date(recent_damage$INCIDENTST,'%m/%d/%Y') # captial Y means four-digit year

```

###Step 1: Filter out 2018
```{r filter_2018}

recent_damage <- recent_damage %>% 
  mutate(YEAR = year(recent_damage$INCIDENTST))

subset_2018 <- recent_damage %>% 
  filter(YEAR == 2018)

```

###Step 2: Figure out how many structures were damaged or destroyed

####Step 2.1: How many structures are there total? How many fires and what are their names?

Each row is a separate structure, and there are 17 types of structures that include things like single family homes, commercial multi-story, commercial single-story, mobile homes, schools, "miscellaneous utility structures", and hospital buildings. By far, the most damaged (in number) were single family single story (12,160 out of 28,388 total).

There were 36 fires.
```{r}

# type of structures damaged in 2018

types_2018 <- as.data.frame(table(subset_2018$STRUCTURET))
total_2018 <- sum(types_2018$Freq) # 28,388 structures

# the fires

fires_2018 <- as.data.frame(table(subset_2018$INCIDENTNA)) # this tells me how many times each fire shows up, which actually means the number of buildings destroyed and damaged! But I need to separate into damaged and destroyed

```

####Step 2.2: Classify structures (aka. rows) as damaged or destroyed

The "DAMAGE" column has XXX options: Affected (1-9%), Destroyed (>50%), Major (26-50%), Minor (10-25%), and No Damage. My gut says to classify Affected, Major, and Minor as "Damaged," Destroyed (duh) as "Destroyed," and to remove No Damage. Then, I should group by Incident Name and count the number of destroyed and damaged structures.

```{r}

# type of damage
damtype_2018 <- as.data.frame(table(subset_2018$DAMAGE))

# classify damage and destruction
damage_2018 <- subset_2018 %>%
  filter(DAMAGE != "No Damage") %>% 
  mutate(DAMAGE_CLASS = ifelse(DAMAGE == "Affected (1-9%)", "Damaged",
                               ifelse(DAMAGE == "Major (26-50%)", "Damaged",
                                      ifelse(DAMAGE == "Minor (10-25%)", "Damaged", "Destroyed"))))

```

```{r}

structures_2018 <- damage_2018 %>% # GOOD ENOUGH
  group_by(INCIDENTNA) %>%
  count(DAMAGE_CLASS)

```

Any other data that I should add for the final CSV that's available in DINS?
```{r}

# Cal Fire Unit, County, Incident Number, Incident start date

units_2018 <- damage_2018 %>% 
  group_by(INCIDENTNA) %>%
  summarize(unit = paste(sort(unique(CALFIREUNI)),collapse=", "))

counties_2018 <- damage_2018 %>% 
  group_by(INCIDENTNA) %>%
  summarize(unit = paste(sort(unique(COUNTY)),collapse=", "))

incinumber_2018 <- damage_2018 %>% 
  group_by(INCIDENTNA) %>%
  summarize(unit = paste(sort(unique(INCIDENTNU)),collapse=", "))

start_2018 <- damage_2018 %>% 
  group_by(INCIDENTNA) %>%
  summarize(unit = paste(sort(unique(INCIDENTST)),collapse=", "))

```


###Step 3: Format like the other Redbook data

I want to add this 2018 damage data to my final CSV (that will eventually be attached to the 2018 fire shapefiles), but first I need to run it through the same wrangling code to add columns, NAs, etc.
```{r}

# get the wrangling code from my other Markdown and run it, then add to NEW/SEPARATE calfire_data and wait until Sam sends attribute table

# data
DINS_2018 <- read_csv("CDF_DINS_damage_2018.csv")

# even though there is only one df, just gonna do the loop
list <- c(DINS_2018)

# the columns to add
cols <- c(Incident_number = NA_real_, Region = NA_real_, Ranger_unit = NA_real_, Ranger_unit2 = NA_real_, National_forest = NA_character_, County = NA_real_, County2 = NA_real_, County3 = NA_real_, Fire_name2 = NA_real_, DPA = NA_real_, DPA2 = NA_real_, CDF_acres_burned = NA_real_, Other_acres_burned = NA_real_, Veg_type = NA_real_, Cause = NA_real_, Destroyed_structures = NA_integer_, Damaged_structures = NA_integer_, Firefighter_fatalities = NA_real_, Civilian_fatalities = NA_real_, Start = NA_real_, Contained = NA_real_)

# the loop
DINS_2018 <- DINS_2018 %>% 
  add_column(!!!cols[!names(cols) %in% names(DINS_2018)]) %>% 
  dplyr::rename(FIRE_NAME = Fire_name)

DINS_2018$FIRE_NAME <- toupper(DINS_2018$FIRE_NAME)

# DONE

# bind to a new calfire_data_2018
library(plyr) # rbind.fill()
calfire_data <- read_csv("calfire_data.csv")
list <- list(DINS_2018, calfire_data)
calfire_data_2018 <- rbind.fill(list)

write_csv(calfire_data_2018, path = "calfire_data_2018.csv")

```

###Step 4: View the new attribute table with 2018 fires
```{r}

fires_2018 <- read_csv(
  "fireshapesforhanna_2018.csv",
  col_types = cols(
    ALARM_DATE = col_date(),
    CONT_DATE = col_date()
  )) # whatever, this doesn't help...

fires_2018$ALARM_DATE <- as.Date(fires_2018$ALARM_DATE)
fires_2018$CONT_DATE <- as.Date(fires_2018$CONT_DATE)

# select fires from 1979 to 2018
fires_2018 <- fires_2018 %>% 
  filter(YEAR_ > 1978)

fires_2017 <- read_csv("fire17_1_1979-2017_damage_final_AMENDED.csv")

# 10,627 entries compared to 10,215 in 1979-2017 -- this makes sense!

write_csv(fires_2018, path = "fire18_1_1979-2018.csv")

```

Now, I am going to copy and paste the 2018 fires from "fire18_1...csv" into "fire17_1...AMENDED.csv".









