---
title: "Manipulating 2018 Cal Fire damage data"
author: "Hanna Buechi"
date: "11/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the data that Steve XXX at Cal Fire sent to me and Dick. It includes information about *each house* (!!!), down to the street number, that was damaged or destroyed in wildfires between 2013 and 2018. I just need the 2018 stuff.

```{r data_packages}

library(tidyverse)
library(lubridate)

recent_damage <- read_csv(
  "DINS2013_2018_attributes.csv",
  col_types = cols(
    REPORTED = col_character(),
    BATTALION = col_character(),
    FIRENAME = col_character()
  ))

recent_damage$INCIDENTST <- as.Date(recent_damage$INCIDENTST,'%m/%d/%Y') # captial Y means four-digit year

```

###Step 1: Filter out 2018
```{r filter_2018}

recent_damage <- recent_damage %>% 
  mutate(YEAR = year(recent_damage$INCIDENTST))

subset_2018 <- recent_damage %>% 
  filter(YEAR == 2018)

```

###Step 2: Figure out how many structures were damaged or destroyed

####Step 2.1: How many structures are there total? How many fires and what are their names?

Each row is a separate structure, and there are 17 types of structures that include things like single family homes, commercial multi-story, commercial single-story, mobile homes, schools, "miscellaneous utility structures", and hospital buildings. By far, the most damaged (in number) were single family single story (12,160 out of 28,388 total).

There were 36 fires.
```{r}

# type of structures damaged in 2018

types_2018 <- as.data.frame(table(subset_2018$STRUCTURET))
total_2018 <- sum(types_2018$Freq) # 28,388 structures

# the fires

fires_2018 <- as.data.frame(table(subset_2018$INCIDENTNA)) # this tells me how many times each fire shows up, which actually means the number of buildings destroyed and damaged! But I need to separate into damaged and destroyed

```

####Step 2.2: Classify structures (aka. rows) as damaged or destroyed

The "DAMAGE" column has XXX options: Affected (1-9%), Destroyed (>50%), Major (26-50%), Minor (10-25%), and No Damage. My gut says to classify Affected, Major, and Minor as "Damaged," Destroyed (duh) as "Destroyed," and to remove No Damage. Then, I should group by Incident Name and count the number of destroyed and damaged structures.

```{r}

# type of damage
damtype_2018 <- as.data.frame(table(subset_2018$DAMAGE))

# classify damage and destruction
damage_2018 <- subset_2018 %>%
  filter(DAMAGE != "No Damage") %>% 
  mutate(DAMAGE_CLASS = ifelse(DAMAGE == "Affected (1-9%)", "Damaged",
                               ifelse(DAMAGE == "Major (26-50%)", "Damaged",
                                      ifelse(DAMAGE == "Minor (10-25%)", "Damaged", "Destroyed"))))

```

```{r}

structures_2018 <- damage_2018 %>% # GOOD ENOUGH
  group_by(INCIDENTNA) %>%
  count(DAMAGE_CLASS)

```

Any other data that I should add for the final CSV that's available in DINS?
```{r}

# Cal Fire Unit, County, Incident Number, Incident start date

units_2018 <- damage_2018 %>% 
  group_by(INCIDENTNA) %>%
  summarize(unit = paste(sort(unique(CALFIREUNI)),collapse=", "))

counties_2018 <- damage_2018 %>% 
  group_by(INCIDENTNA) %>%
  summarize(unit = paste(sort(unique(COUNTY)),collapse=", "))

incinumber_2018 <- damage_2018 %>% 
  group_by(INCIDENTNA) %>%
  summarize(unit = paste(sort(unique(INCIDENTNU)),collapse=", "))

start_2018 <- damage_2018 %>% 
  group_by(INCIDENTNA) %>%
  summarize(unit = paste(sort(unique(INCIDENTST)),collapse=", "))

```


###Step 3: Format like the other Redbook data

I want to add this 2018 damage data to my final CSV (that will eventually be attached to the 2018 fire shapefiles), but first I need to run it through the same wrangling code to add columns, NAs, etc.
```{r}



```

