---
title: "Finding Redbook fires without a match"
author: "Hanna Buechi"
date: "11/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Task: Identify fires from the Redbooks that I could not find a match for in the historical perimeter spatial data.

```{r data_and_packages}

library(tidyverse)
calfire_data <- read_csv("calfire_data.csv",
                         col_types = cols(
                           National_forest = col_character() # specify how to classify National Forest column
                         ))

wildfire_damage <- read_csv(
  "fire17_1_1979-2017_damage_final_AMENDED.csv",
  col_types = cols(
    DPA2_rb = col_character(), # specify how to classify these columns to avoid parsing errors
    Inci_rb = col_character(),
    County2_rb = col_character(),
    CONT_DATE = col_date()
  )
)

```

###Step 1: Concatenate columns in the two datasets.
This is better than matching by fire name because there are duplicate names for different fires. I tried a few combinations. Fire name, year, and ranger unit got me closest to the 508 fires I knew didn't match. Fire name and start date would have been the most specific, but the dates were in different formats between the two datasets and I was too lazy to deal with it.

###Step 2: Subset the fires with Redbook matches from the historical perimeter dataset; check the number of fires in each "new" dataset.
Total Redbook fires: 2,044
Total spatial fires with RB match: 1,536 --> 508 RB fires didn't make it into the final, matched dataset

###Step 3: Convert the concatenated column from wildfire_subset_cat into a list and filter the calfire_cat with that list.
This is just one possibility for accomplishing my task. It makes a subsetted dataframe of calfire_cat of fires that matched. (I can think of a lot of other possibilities now, including removing the spatial data-related columns from wildfire_subset_cat and then performing an anti-join, but ¯\_(ツ)_/¯.)

###Step 4: Compare the "fires that matched" dataframe against calfire_data
My fires anti-join! This makes the dataframe of un-matched Redbook fires that will be my final CSV to upload to TNC_wildfire Box, just for future reference.

```{r redbook_spatial_match}

# Steps 1 & 2: Concatenate columns and subset spatial data
calfire_cat <- calfire_data %>%
  unite(Firename_year, FIRE_NAME, Year, Ranger_unit, remove = F)
  
wildfire_subset_cat <- wildfire_damage %>%
  filter(!is.na(Fire_rb)) %>% # new dataframe with fires that matched, aka. *have* a Redbook fire name =)
  unite(Firename_year, Fire_rb, Year, Ranger_rb, remove = F)

# Step 3
spatial_fires <- c(wildfire_subset_cat$Firename_year) # use this to select from calfire_data

fires_that_match <- calfire_cat %>% 
  filter(Firename_year %in% spatial_fires) # Which "Firename_year" is in "spatial_fires"?

#Step 4
fires_that_dont_match <- anti_join(calfire_cat, fires_that_match)

write_csv(fires_that_dont_match, path = "fires_that_dont_match.csv")

```


