---
title: "Cleaning Redbook data"
author: "Hanna Buechi"
date: "8/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

See Google document for notes and more details: https://docs.google.com/document/d/1FLUtA51ho4nW5CKgt4DMQCBJFIL3fUAzCFjL5bojCOE/edit

###Plan:

1) Standardize column names by hand via in Google Drive: https://drive.google.com/open?id=1uph0UONuYxq34WErMYA9C2FJ8zHDI1FT1ltEbFn2VBY
2) Upload CSV for each year to github folder: EmLab_wildfire
3) Wrangle stuff
4) Write for loop to add new columns (https://stackoverflow.com/questions/45857787/adding-column-if-it-does-not-exist)
5) Bind all CSV's together into a single dataframe (https://stackoverflow.com/questions/2851327/convert-a-list-of-data-frames-into-one-data-frame)

##Wrangling my single dataframe with all of the years:

1) Change "Cal Fire" to "CDF" with recode()
2) Standardize Veg_type
3) Standardize Cause
      Tabulate "Cause" possibilities --> little table function (ESM 244 HW on my laptop) OR count(dataframe, "column heading")

```{r data_packages, message=FALSE}

library(tidyverse)
library(tibble)
library(plyr) # for rbind.fill()
library(readr)
library(lubridate)

rb1979 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1979.csv")
rb1980 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1980.csv")
rb1981 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1981.csv")
rb1982 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1982.csv")
rb1983 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1983.csv")
rb1984 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1984.csv")
rb1985 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1985.csv")
rb1986 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1986.csv")
rb1987 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1987.csv")
rb1988 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1988.csv")
rb1989 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1989.csv")
rb1990 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1990.csv")
rb1991 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1991.csv")
rb1992 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1992.csv")
rb1993 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1993.csv")
rb1994 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1994.csv")
rb1995 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1995.csv")
rb1996 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1996.csv")
rb1997 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1997.csv")
rb1998 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1998.csv")
rb1999 <- read_csv("CDF_fire_damage_1984-2017_V2 - 1999.csv")
rb2000 <- read_csv("CDF_fire_damage_1984-2017_V2 - 2000.csv")
rb2001 <- read_csv("CDF_fire_damage_1984-2017_V2 - 2001.csv")
rb2002 <- read_csv("CDF_fire_damage_1984-2017_V2 - 2002.csv")
rb2003 <- read_csv("CDF_fire_damage_1984-2017_V2 - 2003.csv")
rb2004 <- read_csv("CDF_fire_damage_1984-2017_V2 - 2004.csv")
rb2005 <- read_csv("CDF_fire_damage_1984-2017_V2 - 2005.csv")
rb2006 <- read_csv("CDF_fire_damage_1984-2017_V2 - 2006.csv")
rb2007 <- read_csv("CDF_fire_damage_1984-2017_V2 - 2007.csv")
rb2008a <- read_csv("CDF_fire_damage_1984-2017_V2 - 2008.csv")
rb2008b <- read_csv("CDF_fire_damage_1984-2017_V2 - 2008 non-CDF.csv")
rb2009a <- read_csv("CDF_fire_damage_1984-2017_V2 - 2009.csv")
rb2009b <- read_csv("CDF_fire_damage_1984-2017_V2 - 2009 non-CDF.csv")
rb2010 <- read_csv("CDF_fire_damage_1984-2017_V2 - 2010.csv")
rb2011 <- read_csv("CDF_fire_damage_1984-2017_V2 - 2011.csv")
rb2012 <- read_csv("CDF_fire_damage_1984-2017_V2 - 2012.csv")
rb2013 <- read_csv("CDF_fire_damage_1984-2017_V2 - 2013.csv")
rb2014 <- read_csv("CDF_fire_damage_1984-2017_V2 - 2014.csv")
rb2015 <- read_csv("CDF_fire_damage_1984-2017_V2 - 2015.csv")
rb2016 <- read_csv("CDF_fire_damage_1984-2017_V2 - 2016.csv")
rb2017 <- read_csv("CDF_fire_damage_1984-2017_V2 - 2017.csv")

```

###Add columns where needed
```{r add_column_loop}

# I need to add columns to most, if not all, of the Cal Fire dataframes so that they can be bound together into a single dataframe. Thanks, Stack Overflow, for the code!

list <- list(redbook_1979, redbook_1980, redbook_1981, redbook_1982, redbook_1983, redbook_1984, redbook_1985, redbook_1986, redbook_1987, redbook_1988, redbook_1989, redbook_1990, redbook_1991, redbook_1992, redbook_1993, redbook_1994, redbook_1995, redbook_1996, redbook_1997, redbook_1998, redbook_1999, redbook_2000, redbook_2001, redbook_2002, redbook_2003, redbook_2004, redbook_2005, redbook_2006, redbook_2007, redbook_2008a, redbook_2008b, redbook_2009a, redbook_2009b, redbook_2010, redbook_2011, redbook_2012, redbook_2013, redbook_2014, redbook_2015, redbook_2016, redbook_2017) # list of dfs for the forloop

cols <- c(Incident_number = NA_real_, Region = NA_real_, Ranger_unit = NA_real_, Ranger_unit2 = NA_real_, National_forest = NA_character_, County = NA_real_, County2 = NA_real_, County3 = NA_real_, Fire_name2 = NA_real_, DPA = NA_real_, DPA2 = NA_real_, CDF_acres_burned = NA_real_, Other_acres_burned = NA_real_, Veg_type = NA_real_, Cause = NA_real_, Destroyed_structures = NA_integer_, Damaged_structures = NA_integer_, Firefighter_fatalities = NA_real_, Civilian_fatalities = NA_real_, Start = NA_real_, Contained = NA_real_) 
# vector of column headings to be populated with NAs if they are being added

for (i in 1:length(list)){
 list[[i]] <- add_column(list[[i]], !!!cols[!names(cols) %in% names(list[[i]])])
} # looped-de-loop!

calfire_data <- rbind.fill(list) # binds the list of dataframes together

```

###Wrangle stuff
```{r check_classes_of_columns}

#lapply(calfire_data, class) # check the class of each column in the dataframe and change as needed
# belows are the changes that needed to be made

calfire_data$Cause <- as.character(calfire_data$Cause)
calfire_data$CDF_acres_burned <- as.numeric(calfire_data$CDF_acres_burned)
calfire_data$Other_acres_burned <- as.numeric(calfire_data$Other_acres_burned)
calfire_data$Destroyed_structures <- as.numeric(calfire_data$Destroyed_structures)
calfire_data$Damaged_structures <- as.numeric(calfire_data$Damaged_structures) 
calfire_data$Firefighter_fatalities <- as.numeric(calfire_data$Firefighter_fatalities)
calfire_data$Civilian_fatalities <- as.numeric(calfire_data$Civilian_fatalities)
#calfire_data$Start <- as.Date(calfire_data$Start)
#calfire_data$Contained <- as.Date(calfire_data$Contained)

# damaged, firefighter, and civilian have N/As
# damaged has "Yes" for Devils 1994, 100's for Wheel 1993

```

###Standardize vegetation type, cause
```{r standardize_veg_CDF}

# check out all of the values for Veg_type and DPA, standardize as needed. I'm sure this could be made into shorter code, but I didn't now how. See Google Drive for the veg type codes: https://drive.google.com/open?id=1XVDiNHeqq2jxiQqYmb657MMfjK9UIYU6PorzYV_QubY

#unique(calfire_data$Veg_type) # shows me all of the possibilities so that I can copy and paste...
#unique(calfire_data$DPA2)

calfire_data <- calfire_data %>% 
  mutate(DPA=recode(DPA, "Cal Fire"="CDF", "CalFire"="CDF")) %>% 
  mutate(Veg_type=recode(Veg_type, "CONIFERS"="C")) %>%
  mutate(Veg_type=recode(Veg_type, "G/AGPROD"="GA")) %>%
  mutate(Veg_type=recode(Veg_type, "Timber"="T")) %>%
  mutate(Veg_type=recode(Veg_type, "Brush"="B")) %>%
  mutate(Veg_type=recode(Veg_type, "Grass, Brush"="GB")) %>%
  mutate(Veg_type=recode(Veg_type, "Brush, Grass"="BG")) %>%
  mutate(Veg_type=recode(Veg_type, "Woodland"="W")) %>%
  mutate(Veg_type=recode(Veg_type, "Brush, Grass, Timber"="BGT")) %>%
  mutate(Veg_type=recode(Veg_type, "Grass"="G")) %>%
  mutate(Veg_type=recode(Veg_type, "Timber, Grass"="TG")) %>%
  mutate(Veg_type=recode(Veg_type, "Timber, Brush, Grass"="TBG")) %>%
  mutate(Veg_type=recode(Veg_type, "Unknown"="U")) %>%
  mutate(Veg_type=recode(Veg_type, "Timber, Woodland"="TW")) %>%
  mutate(Veg_type=recode(Veg_type, "Timber, Brush"="TB")) %>%
  mutate(Veg_type=recode(Veg_type, "Brush, Timber"="BT")) %>%
  mutate(Veg_type=recode(Veg_type, "Timer, Brush"="TB")) %>% 
  mutate(Veg_type=recode(Veg_type, "BGAP"="BGA")) %>%
  mutate(Veg_type=recode(Veg_type, "WBGAP"="WBGA")) %>%
  mutate(Veg_type=recode(Veg_type, "WAP"="WA")) %>%
  mutate(Veg_type=recode(Veg_type, "TWAP"="TWA")) %>%
  mutate(Veg_type=recode(Veg_type, "BAP"="BA")) %>%
  mutate(Veg_type=recode(Veg_type, "TGWAPB"="TGWAB")) %>%
  mutate(Veg_type=recode(Veg_type, "GAPB"="GAB"))

```

```{r standardize_cause}

# same notes as above, but there are some causes that I haven't standardized yet. See list at the bottom of this chunk...

#unique(calfire_data$Cause)

calfire_data <- calfire_data %>% 
  mutate(Cause=recode(Cause, "Equip. Use"="Equipment use")) %>% 
  mutate(Cause=recode(Cause, "N/A"="NA")) %>%   
  mutate(Cause=recode(Cause, "Cmpr."="Campfire")) %>%
  mutate(Cause=recode(Cause, "Smkr."="Smoking")) %>% 
  mutate(Cause=recode(Cause, "Misc."="Misc")) %>% 
  mutate(Cause=recode(Cause, "Elec. Power"="Electrical power")) %>% 
  mutate(Cause=recode(Cause, "P/W/F"="Playing with fire")) %>% 
  mutate(Cause=recode(Cause, "Ltng."="Lightning")) %>% 
  mutate(Cause=recode(Cause, "Smkg."="Smoking")) %>% 
  mutate(Cause=recode(Cause, "Play./W/Fire"="Playing with fire")) %>% 
  mutate(Cause=recode(Cause, "Playing/W Fire"="Playing with fire")) %>% 
  mutate(Cause=recode(Cause, "Miscellaneous"="Misc")) %>% 
  mutate(Cause=recode(Cause, "Equip."="Equipment use")) %>% 
  mutate(Cause=recode(Cause, "Play w/fire"="Playing with fire")) %>% 
  mutate(Cause=recode(Cause, "Power"="Electrical power")) %>%
  mutate(Cause=recode(Cause, "Undet."="Undetermined")) %>%
  mutate(Cause=recode(Cause, "Smoker"="Smoking")) %>%
  mutate(Cause=recode(Cause, "Play/W/Fire"="Playing with fire")) %>%
  mutate(Cause=recode(Cause, "Arcing Pwerline"="Powerlines")) %>%
  mutate(Cause=recode(Cause, "Powerline"="Powerlines")) %>%
  mutate(Cause=recode(Cause, "Electrical Pwr"="Electrical power")) %>%
  mutate(Cause=recode(Cause, "Exposure Fire"="Exposure")) %>%
  mutate(Cause=recode(Cause, "Equipment"="Equipment use")) %>%
  mutate(Cause=recode(Cause, "Elec. Pow."="Electrical power")) %>%
  mutate(Cause=recode(Cause, "Munition"="Shooting")) %>%
  mutate(Cause=recode(Cause, "PWF"="Playing with fire")) %>%
  mutate(Cause=recode(Cause, "Elect. Power"="Electrical power")) %>%
  mutate(Cause=recode(Cause, "Under Inv."="Under investigation")) %>%
  mutate(Cause=recode(Cause, "Lawn Mower"="Equipment use")) %>%
  mutate(Cause=recode(Cause, "Under Invest."="Under investigation")) %>%
  mutate(Cause=recode(Cause, "Shooter"="Shooting")) %>%
  mutate(Cause=recode(Cause, "Welding"="Equipment use")) %>%
  mutate(Cause=recode(Cause, "P-W-F"="Playing with fire")) %>%
  mutate(Cause=recode(Cause, "Inidentified"="Unidentified")) %>%
  mutate(Cause=recode(Cause, "Miscellaneous Human"="Misc")) %>%
  mutate(Cause=recode(Cause, "Other Cause"="Other")) %>%
  mutate(Cause=recode(Cause, "Electrical"="Electrical power")) %>%
  mutate(Cause=recode(Cause, "UI"="Unidentified")) %>%
  mutate(Cause=recode(Cause, "LIghtning"="Lightning")) %>%
  mutate(Cause=recode(Cause, "Electrical Power"="Electrical power")) %>%
  mutate(Cause=recode(Cause, "Playing with Fire"="Playing with fire"))

# Rekindle debris? Equip. RR? Exposure? VMP Burn? UL? Rekindle? Control Burn? Hostile Fire? Escape Control? Debris Burning? Human? Structure? Debris, veg burn? Incediary? Open Outdoors? Miscellaneous Human? Misuse of Fire?

write_csv(calfire_data, path = "calfire_data.csv")

```











